---
title: "Analyses MMQ: Translation, Adaptation and Psychometric Properties in a Chilean Population (2023)"
author: "Edmundo Kronmuller & Cristian Brotfeld"
date: "2023-07-24"
output:
  html_document: default
  pdf_document: default
---

### Set up, load packages and data

```{r setup, include=TRUE, message=FALSE, results=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
pacman::p_load(tidyverse, lavaan, psych, ltm, labelled, apaTables,arsenal,Hmisc,writexl,semTools)


# Read data
datos <- read_rds("datos.rds")
```

### Sample description of participants with no missing data on the MMQ

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}
naniar::miss_var_summary(datos[,c(3:6,7:39)])
socio.mmq <- datos[,c(3:6,7:39)]
socio.mmq$edad <- ifelse(socio.mmq$edad==0, mean(socio.mmq$edad, na.rm=TRUE),socio.mmq$edad)
socio.mmq <- na.omit(socio.mmq)
socio.mmq <- socio.mmq[,c(1:4)]
socio.mmq <- unlabelled(socio.mmq)
names(socio.mmq)

tab.sociodemo.mmq <- tableby(~edad+genero+educacion, data = socio.mmq)

as.data.frame(summary(tab.sociodemo.mmq, text=NULL))

write.table(file="sociodemo.mmq.text", tab.sociodemo.mmq <- as.data.frame(
  summary(tab.sociodemo.mmq, text=NULL)))

```


### Descriptives by item: Frequencies

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# dataframes with labels 
datos_un.mmq = unlabelled(datos[,7:39])
nombres.mmq = labelled::var_label(datos_un.mmq)
nombres.mmq = unlist(nombres.mmq)
items.un.mmq = datos_un.mmq %>% dplyr::select(1:33)
items.un.mmq = na.omit(items.un.mmq) 

# descriptives by item
nombres.mmq <- str_sub(nombres.mmq, start = 1, end = 100)
colnames(items.un.mmq) <- nombres.mmq

desc.resp.mmq.long.pre <-
  items.un.mmq %>% 
  gather(key=item, value = respuesta) %>% 
  mutate(respuesta = factor(respuesta, 
                            levels = c("Totalmente en desacuerdo", "En desacuerdo",
                                       "Ni de acuerdo ni en desacuerdo", "De acuerdo",
                                       "Muy de acuerdo")))
desc.resp.mmq.long <-   
  desc.resp.mmq.long.pre %>%
  dplyr::count(item,respuesta) %>% group_by(item) %>% 
  mutate(porc = round(n/sum(n),3)*100)

desc.resp.mmq.wide <- 
  desc.resp.mmq.long %>% 
    dplyr::select(item, respuesta, porc) %>%
    pivot_wider(names_from = respuesta, 
              values_from = porc)

print(desc.resp.mmq.wide)


```



### Descriptives per item: Mean, standard deviation, ...

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}
# dataframe for descriptives and cfa
mmq = datos[,7:39]
mmq = na.omit(mmq)
# descriptives
d1 <- psych::describe(mmq, type = 2, skew=TRUE,range=TRUE)
reactivo.mmq <- data.frame(nombres.mmq)
descriptivos.items <- cbind(reactivo.mmq, d1)

print(descriptivos.items)

```



### Confirmatiory Factor Analysis on original scale (Gori et al., 2021)

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition
model.mmq.or <- '
refv =~ q0007_0001 + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017 + q0007_0018 + q0007_0031 + q0007_0032

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025 + q0007_0026 + 
  q0007_0030

relatv =~ q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'

# fit 
mmq.fit.or <-cfa(model.mmq.or, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(mmq.fit.or, fit.measures = TRUE)

```


### Modification indices analysis
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# modification indices
mi.or <- modindices(mmq.fit.or, sort=TRUE) 
mi.or.1 <- mi.or[mi.or$op == "=~",]

print(mi.or.1)

```

### CFA on modified scale
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition
model.mmq.mod <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017  

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028 

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'

mmq.fit.mod <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(mmq.fit.mod, fit.measures = TRUE)

#summary(mmq.fit, standardized=TRUE)
#summary(mmq.fit.mod, fit.measures = TRUE)

#mmq.fit.mod <-cfa_fit_plot(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

#fittable.mod <-nice_fit(mmq.fit.mod, nice_table = TRUE)
#flextable::save_as_docx(fittable.mod, path = "fittable_mod.docx")

#loads.mod <- data.frame(inspect(mmq.fit.mod,what="std")$lambda)
#loads.mod <- round(loads.mod,3)
#loads.mod$names <- rownames(loads.mod)

#write.csv(loads.mod, "loads.mod.csv", row.names = F)

parameterestimates(mmq.fit.mod) %>% filter(op == "~~")


```



Modelos alternativos:


## Bifactor:


```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition
model.mmq.mod_bifactor <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017  

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028 

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023

global =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017 + q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025 +
  q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028 +
  q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033 +
  q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029 +
  q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
  
refv~~0*egov
refv~~0*relatv
refv~~0*reldiv
refv~~0*distrv
refv~~0*emdisv
egov~~0*relatv
egov~~0*reldiv
egov~~0*distrv
egov~~0*emdisv
relatv~~0*reldiv
relatv~~0*distrv
relatv~~0*emdisv
reldiv~~0*distrv
reldiv~~0*emdisv
distrv~~0*emdisv
global ~~ 0*refv
global ~~ 0*egov
global ~~ 0*relatv
global ~~ 0*reldiv
global ~~ 0*emdisv
global ~~ 0*distrv
'

mmq.fit.mod_bifactor <-cfa(model.mmq.mod_bifactor, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(mmq.fit.mod_bifactor, fit.measures = TRUE)
fitMeasures(mmq.fit.mod_bifactor, c("chisq", "df", "pvalue", "cfi", "tli","rmsea","srmr"))
#summary(mmq.fit, standardized=TRUE)
#summary(mmq.fit.mod, fit.measures = TRUE)

#mmq.fit.mod <-cfa_fit_plot(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

#fittable.mod <-nice_fit(mmq.fit.mod, nice_table = TRUE)
#flextable::save_as_docx(fittable.mod, path = "fittable_mod.docx")

#loads.mod <- data.frame(inspect(mmq.fit.mod,what="std")$lambda)
#loads.mod <- round(loads.mod,3)
#loads.mod$names <- rownames(loads.mod)

#write.csv(loads.mod, "loads.mod.csv", row.names = F)
1460.47/322

```



## Un solo factor:


```{r}
one_factor = 'global =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017 + q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025 +
  q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028 +
  q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033 +
  q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029 +
  q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023'

one_factor <-cfa(one_factor, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(one_factor, fit.measures = TRUE)
fitMeasures(one_factor, c("chisq", "df", "pvalue", "cfi", "tli","rmsea","srmr"))
4649.834/350.000
```


## Modelo jerárquico

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition
model.mmq.hierarchical <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017  

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028 

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023


global =~ refv + egov + relatv + reldiv + distrv + emdisv

'

mmq.fit.mod_hierarchical <-cfa(model.mmq.hierarchical, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(mmq.fit.mod_hierarchical, fit.measures = TRUE)
fitMeasures(mmq.fit.mod_hierarchical, c("chisq", "df", "pvalue", "cfi", "tli","rmsea","srmr"))
#summary(mmq.fit, standardized=TRUE)
#summary(mmq.fit.mod, fit.measures = TRUE)

#mmq.fit.mod <-cfa_fit_plot(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

#fittable.mod <-nice_fit(mmq.fit.mod, nice_table = TRUE)
#flextable::save_as_docx(fittable.mod, path = "fittable_mod.docx")

#loads.mod <- data.frame(inspect(mmq.fit.mod,what="std")$lambda)
#loads.mod <- round(loads.mod,3)
#loads.mod$names <- rownames(loads.mod)

#write.csv(loads.mod, "loads.mod.csv", row.names = F)

1664.929/344
```




## Invarianza:

```{r}
mmq2 = datos[,c(4,7:39)]
mmq2 = na.omit(mmq2)
table(mmq2$genero)
```


```{r}
mmq2$genero = factor(mmq2$genero)
mmq2 =  mmq2 %>% filter(genero %in% c("1","2"))
```

### CFA on modified scale
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition 
# sacar ítems porque faltan valores: q0007_0014, q0007_0028,q0007_0008, q0007_0017
model.mmq.mod <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0016

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005  

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'

mmq.fit.mod_conf <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")
mmq.fit.mod_tresh <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds"))
mmq.fit.mod_tresh_loadings <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds","loadings"))
mmq.fit.mod_tresh_loadings_intercepts <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds","loadings","intercepts"))
mmq.fit.mod_tresh_loadings_intercepts_residuals <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds","loadings","intercepts","residuals","residuals"))


summary(mmq.fit.mod_conf, fit.measures = TRUE)
summary(mmq.fit.mod_tresh, fit.measures = TRUE)
summary(mmq.fit.mod_tresh_loadings, fit.measures = TRUE)
summary(mmq.fit.mod_tresh_loadings_intercepts, fit.measures = TRUE)

#summary(mmq.fit, standardized=TRUE)
#summary(mmq.fit.mod, fit.measures = TRUE)

#mmq.fit.mod <-cfa_fit_plot(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

#fittable.mod <-nice_fit(mmq.fit.mod, nice_table = TRUE)
#flextable::save_as_docx(fittable.mod, path = "fittable_mod.docx")

#loads.mod <- data.frame(inspect(mmq.fit.mod,what="std")$lambda)
#loads.mod <- round(loads.mod,3)
#loads.mod$names <- rownames(loads.mod)

#write.csv(loads.mod, "loads.mod.csv", row.names = F)

```

```{r}
anova(mmq.fit.mod_conf, mmq.fit.mod_tresh, mmq.fit.mod_tresh_loadings, mmq.fit.mod_tresh_loadings_intercepts)
```





### Descriptives on modified scale and subscales and correlations
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# means modified scales
mmq$refle.mod <- rowMeans(mmq[,c( 1,6,8,10,16,17)])
mmq$egost.mod = rowMeans(mmq[,c(11,22,24,25)])
mmq$relatt..mod = rowMeans(mmq[,c(4,5,14,21,28)])
mmq$reldis..mod = rowMeans(mmq[,c(9,12,15,27,33)])
mmq$distr.mod = rowMeans(mmq[,c(13,19,20,29)])
mmq$emodis.mod = rowMeans(mmq[,c(2,3,7,23)])

mmq$mmqPositive <- rowMeans(mmq[,c(34,35,36)])
mmq$mmqNegative <- rowMeans(mmq[,c(37,38,39)])


# test Shapiro
s1 <- apply(mmq[,34:41], 2, shapiro.test)[[1]][[2]]
s2 <- apply(mmq[,34:41], 2, shapiro.test)[[2]][[2]]
s3 <- apply(mmq[,34:41], 2, shapiro.test)[[3]][[2]]
s4 <- apply(mmq[,34:41], 2, shapiro.test)[[4]][[2]]
s5 <- apply(mmq[,34:41], 2, shapiro.test)[[5]][[2]]
s6 <- apply(mmq[,34:41], 2, shapiro.test)[[6]][[2]]
s7 <- apply(mmq[,34:41], 2, shapiro.test)[[7]][[2]]
s8 <- apply(mmq[,34:41], 2, shapiro.test)[[8]][[2]]

shapiro.mod <- unlist(c(s1,s2,s3,s4,s5,s6,s7,s8))


# Cronbach's alpha
a1 <- alpha(mmq[,c( 1,6,8,10,16,17)])$total[2] #Reflexivity
a2 <-alpha(mmq[,c(11,22,24,25)])$total[2] #Ego-Strenght
a3 <-alpha(mmq[,c(4,5,14,21,28)])$total[2]#Relational Attunement
a4 <-alpha(mmq[,c(9,12,15,27,33)])$total[2]# Relational Discomfort
a5 <-alpha(mmq[,c(13,19,20,29)])$total[2]#Distrust
a6 <-alpha(mmq[,c(2,3,7,23)])$total[2]#Emotional Discontrol
apos <- alpha(mmq[,c( 1,6,8,10,16,17, 11,22,24,25,4,5,14,21,28)])$total[2]
aneg <- alpha(mmq[,c(9,12,15,27,33,13,19,20,29,2,3,7,23)])$total[2]

alphas.mod <- unlist(c(a1,a2,a3,a4,a5,a6,apos,aneg))


d2 <- psych::describe(mmq[,34:41], type = 2, skew=TRUE,range=TRUE)
descrip.scales <- cbind(d2,shapiro.mod,alphas.mod)
descrip.scales$scales <- rownames(descrip.scales)

print(descrip.scales)


# correlation among scales
rcorr(as.matrix(mmq[,34:39]))


```



### Convergent and divergent analysis: correlation matrix
```{r eval = TRUE, message=FALSE, results=FALSE, warning=FALSE}

# create dataframe with all subscales
# MMQ
scales_mmq = datos %>% dplyr::select(starts_with("q0007"))

scales_mmq$mmq_ref <- rowMeans(scales_mmq[,c( 1,6,8,10,16,17)])
scales_mmq$mmq_ego = rowMeans(scales_mmq[,c(11,22,24,25)])
scales_mmq$mmq_relat = rowMeans(scales_mmq[,c(4,5,14,21,28)])
scales_mmq$mmq_reldi = rowMeans(scales_mmq[,c(9,12,15,27,33)])
scales_mmq$mmq_dist = rowMeans(scales_mmq[,c(13,19,20,29)])
scales_mmq$mmq_emdis = rowMeans(scales_mmq[,c(2,3,7,23)])

scales_mmq$mmqPositive <- rowMeans(scales_mmq[,c(34,35,36)])
scales_mmq$mmqNegative <- rowMeans(scales_mmq[,c(37,38,39)])

scales_mmq <- 
  scales_mmq  %>%
    dplyr::select(mmq_ref, mmq_ego, mmq_relat, mmq_reldi, mmq_dist, mmq_emdis,
                  mmqPositive,mmqNegative)


#Empathy Quotient [EQ]: ec: cognitive empathy; hs: social skills; re: emotional reactivity
scales_eq <- datos %>% dplyr::select(starts_with("q0009"))

scales_eq <- scales_eq %>% 
  dplyr::mutate(across(.cols = c(4,8,12,14,15,27,32,35,46,48,49,50), .fns = function(x){x=5-x}))

scales_eq=scales_eq %>%
  dplyr::mutate(across(.cols = everything(), .fns = function(x){recode(x, `1`= 2, `2`=1, `3`=0, `4`=0)}))

scales_eq$eq_ec <- rowMeans(scales_eq[,c(1,19,25,26,36,41,43,44,52,54,55,58,60)])
scales_eq$eq_hs <- rowMeans(scales_eq[,c(4,8,12,14,21,35,57)])
scales_eq$eq_re <- rowMeans(scales_eq[,c(15,27,32,42,46,48,49,50,59)])

scales_eq$eqTotal <- rowMeans(scales_eq[,c(61,62,63)])

scales_eq <- 
  scales_eq  %>%
    dplyr::select(eq_ec, eq_hs, eq_re, eqTotal)


# Interpersonal Reactivity Index (IRI). fs: fantasy scale; pt: perspective taking; ec :empathic concern; pd: personal distress 
scales_iri = datos %>% dplyr::select(starts_with("q0010"))
scales_iri$fs = NULL
scales_iri = scales_iri %>% dplyr::mutate(across(.cols = c(3,4,7,12,13,14,15,18,19), .fns = function(x){6-x}))
scales_iri = scales_iri %>% dplyr::mutate(across(.cols = everything() , .fns = function(x){x-1}))

scales_iri$iri_fs = rowMeans(scales_iri[,c(1,5,7,12,16,23,26)])
scales_iri$iri_pd = rowMeans(scales_iri[,c(6,10,13,17,19,24,27)])
scales_iri$iri_pt = rowMeans(scales_iri[,c(3,8,11,15,21,25,28)])
scales_iri$iri_ec = rowMeans(scales_iri[,c(2,4,9,14,18,20,22)])

scales_iri$iriTotal <- rowMeans(scales_iri[,c(29,30,31,32)])

scales_iri <- 
  scales_iri  %>%
    dplyr::select(iri_fs, iri_pd, iri_pt, iri_ec, iriTotal)


# Buss Perry Aggression Questionnaire (BPAQ): af: physical agression; av: verbal agression; ang: anger; host: hostility
scales_bpaq = datos %>% dplyr::select(starts_with("q0011"))

scales_bpaq$bpaq_af = rowMeans(scales_bpaq[,c(1,5,9,17,21,27,29)])
scales_bpaq$bpaq_av = rowMeans(scales_bpaq[,c(2,6,14,18)])
scales_bpaq$bpaq_ang = rowMeans(scales_bpaq[,c(3,7,11,22)])
scales_bpaq$bpaq_host = rowMeans(scales_bpaq[,c(4,16,20,26,28)])
scales_bpaq$bpaqTotal = rowMeans(scales_bpaq[,c(30,31,32,33)])

scales_bpaq <- 
  scales_bpaq  %>%
    dplyr::select(bpaq_af, bpaq_av, bpaq_ang, bpaq_host,bpaqTotal)


datos_cd <- cbind(
scales_mmq, scales_eq, scales_iri, scales_bpaq)


# correlation matrix
rcorr(as.matrix(datos_cd))
```



# Correlation using latent variables:


```{r}
mmq3 = datos[,7:39]
mmq3 = data.frame(cbind(mmq3,datos_cd))
```




```{r}
covariables_corr = colnames(mmq3[,42:55])
lista_variables = c("refv","egov","relatv","reldiv","distrv","emdisv")
base_correlaciones_latentes = data.frame()
base_significancia_latentes = data.frame()

for(i in covariables_corr){
  var = i
  model.mmq.mod_regresion <- '
  refv =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017  
  
  egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025
  
  relatv =~ q0007_0005 + q0007_0014 + q0007_0021 
  
  reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033
  
  distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029
  
  emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023;
  '
  
  
  
  regresion = str_c(lista_variables, "~", var,collapse = ";")
  model.mmq.mod_regresion = str_c(model.mmq.mod_regresion,regresion)
  model.mmq.mod_regresion
  
  mmq.fit.mod.regression <-sem(model.mmq.mod_regresion, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq3)
  dat = parameterestimates(mmq.fit.mod.regression,standardized = TRUE) %>% filter(op == "~") %>% dplyr::select(lhs, std.all)
  dat$std.all = round(dat$std.all,2)
  dat = dat %>% spread(key = lhs, value = std.all)
  dat$cov = var
  base_correlaciones_latentes = base_correlaciones_latentes %>% bind_rows(dat)
  
  
  dat_p = parameterestimates(mmq.fit.mod.regression,standardized = TRUE) %>% filter(op == "~") %>% dplyr::select(lhs,pvalue)
  dat_p$pvalue = round(dat_p$pvalue,3)
  dat_p = dat_p %>% spread(key = lhs, value = pvalue)
  dat_p$cov = var
  base_significancia_latentes = base_significancia_latentes %>% bind_rows(dat_p)
}

base_correlaciones_latentes
base_significancia_latentes
```


```{r}
write_xlsx(base_correlaciones_latentes, path = "correlaciones.xlsx")
write_xlsx(base_significancia_latentes, path = "pvalues.xlsx")
```



## Invarianza:



```{r}
mmq2 = datos[,c(4,7:39)]
mmq2 = na.omit(mmq2)
table(mmq2$genero)
```


```{r}
mmq2$genero = factor(mmq2$genero)
mmq2 =  mmq2 %>% filter(genero %in% c("1","2"))
```

### CFA on modified scale
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition 
# sacar ítems porque faltan valores: q0007_0014, q0007_0028,q0007_0008, q0007_0017
model.mmq.mod <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0016

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005  

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'

mmq.fit.mod_conf <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")
mmq.fit.mod_tresh <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds"))
mmq.fit.mod_tresh_loadings <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds","loadings"))
mmq.fit.mod_tresh_loadings_intercepts <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds","loadings","intercepts"))
mmq.fit.mod_tresh_loadings_intercepts_residuals <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero",group.equal = c("thresholds","loadings","intercepts","residuals","residuals"))


summary(mmq.fit.mod_conf, fit.measures = TRUE)
summary(mmq.fit.mod_tresh, fit.measures = TRUE)
summary(mmq.fit.mod_tresh_loadings, fit.measures = TRUE)
summary(mmq.fit.mod_tresh_loadings_intercepts, fit.measures = TRUE)

#summary(mmq.fit, standardized=TRUE)
#summary(mmq.fit.mod, fit.measures = TRUE)

#mmq.fit.mod <-cfa_fit_plot(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

#fittable.mod <-nice_fit(mmq.fit.mod, nice_table = TRUE)
#flextable::save_as_docx(fittable.mod, path = "fittable_mod.docx")

#loads.mod <- data.frame(inspect(mmq.fit.mod,what="std")$lambda)
#loads.mod <- round(loads.mod,3)
#loads.mod$names <- rownames(loads.mod)

#write.csv(loads.mod, "loads.mod.csv", row.names = F)

```

```{r}
anova(mmq.fit.mod_conf, mmq.fit.mod_tresh, mmq.fit.mod_tresh_loadings, mmq.fit.mod_tresh_loadings_intercepts)
```




```{r}
model.mmq.mod <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0016

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005  

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'
```


```{r}
script_thresholds <- measEq.syntax(configural.model = model.mmq.mod,
data = mmq2, ordered = c(names(mmq2)),
parameterization = "delta",
ID.fac = "std.lv",
ID.cat = "Wu.Estabrook.2016",
group = "genero",
group.equal = c("thresholds"))
script_thresholds <- as.character(script_thresholds)


## 

script_thresholds_loadings <- measEq.syntax(configural.model = model.mmq.mod,
data = mmq2, ordered = c(names(mmq2)),
parameterization = "delta",
ID.fac = "std.lv",
ID.cat = "Wu.Estabrook.2016",
group = "genero",
group.equal = c("thresholds","loadings"))
script_thresholds_loadings <- as.character(script_thresholds_loadings)


script_thresholds_loadings_intercept <- measEq.syntax(configural.model = model.mmq.mod,
data = mmq2, ordered = c(names(mmq2)),
parameterization = "delta",
ID.fac = "std.lv",
ID.cat = "Wu.Estabrook.2016",
group = "genero",
group.equal = c("thresholds","loadings","intercepts"))
script_thresholds_loadings_intercept <- as.character(script_thresholds_loadings_intercept)

##

script_thresholds_loadings_intercept_residuals <- measEq.syntax(configural.model = model.mmq.mod,
data = mmq2, ordered = c(names(mmq2)),
parameterization = "theta",
ID.fac = "std.lv",
ID.cat = "Wu.Estabrook.2016",
group = "genero",
group.equal = c("thresholds","loadings","intercepts","residuals"))
script_thresholds_loadings_intercept_residuals <- as.character(script_thresholds_loadings_intercept_residuals)


```

```{r}
mmq.fit.mod <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")
mmq.fit.mod_tresh <-cfa(script_thresholds, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")
mmq.fit.mod_tresh_loadings <-cfa(script_thresholds_loadings, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")
mmq.fit.mod_tresh_loadings_intercept <-cfa(script_thresholds_loadings_intercept, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")

mmq.fit.mod_tresh_loadings_intercept_residuals <-cfa(script_thresholds_loadings_intercept_residuals, estimator="WLSMV", ordered=names(mmq2), std.lv=TRUE, data=mmq2,group = "genero")
```


```{r}
anova(mmq.fit.mod,mmq.fit.mod_tresh,mmq.fit.mod_tresh_loadings, mmq.fit.mod_tresh_loadings_intercept,mmq.fit.mod_tresh_loadings_intercept_residuals)
```

```{r}
fitmeasures(mmq.fit.mod,fit.measures = c("cfi","tli","RMSEA","SRMR"))
fitmeasures(mmq.fit.mod_tresh,fit.measures = c("cfi","tli","RMSEA","SRMR"))
fitmeasures(mmq.fit.mod_tresh_loadings,fit.measures = c("cfi","tli","RMSEA","SRMR"))
fitmeasures(mmq.fit.mod_tresh_loadings_intercept,fit.measures = c("cfi","tli","RMSEA","SRMR"))
fitmeasures(mmq.fit.mod_tresh_loadings_intercept_residuals,fit.measures = c("cfi","tli","RMSEA","SRMR"))
```



```{r}
datos %>% ggplot(aes(x = edad)) + geom_density(fill="firebrick") +
  geom_vline(xintercept = median(datos$edad,na.rm=T)) + theme_bw()
ggsave(filename = "edad.jpg", units = "cm", width = 10, height = 8)
```


