---
title: "Analyses MMQ: Translation, Adaptation and Psychometric Properties in a Chilean Population (2023)"
author: "XXX"
date: "2023-07-24"
output:
  html_document: default
  pdf_document: default
---

### Set up, load packages and data

```{r setup, include=TRUE, message=FALSE, results=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(tidyverse)
library(lavaan)
library(psych)
library(ltm)
library(labelled)
library(apaTables)
library(arsenal)
library(Hmisc)

# Read data
datos <- read_rds("datos.rds")

```

### Sample description of participants with no missing data on the MMQ
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}
naniar::miss_var_summary(datos[,c(3:6,7:39)])
socio.mmq <- datos[,c(3:6,7:39)]
socio.mmq$edad <- ifelse(socio.mmq$edad==0, mean(socio.mmq$edad, na.rm=TRUE),socio.mmq$edad)
socio.mmq <- na.omit(socio.mmq)
socio.mmq <- socio.mmq[,c(1:4)]
socio.mmq <- unlabelled(socio.mmq)
names(socio.mmq)

tab.sociodemo.mmq <- tableby(~edad+genero+educacion, data = socio.mmq)

as.data.frame(summary(tab.sociodemo.mmq, text=NULL))

write.table(file="sociodemo.mmq.text", tab.sociodemo.mmq <- as.data.frame(
  summary(tab.sociodemo.mmq, text=NULL)))

```


### Descriptives by item: Frequencies

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# dataframes with labels 
datos_un.mmq = unlabelled(datos[,7:39])
nombres.mmq = labelled::var_label(datos_un.mmq)
nombres.mmq = unlist(nombres.mmq)
items.un.mmq = datos_un.mmq %>% dplyr::select(1:33)
items.un.mmq = na.omit(items.un.mmq) 

# descriptives by item
nombres.mmq <- str_sub(nombres.mmq, start = 1, end = 100)
colnames(items.un.mmq) <- nombres.mmq

desc.resp.mmq.long.pre <-
  items.un.mmq %>% 
  gather(key=item, value = respuesta) %>% 
  mutate(respuesta = factor(respuesta, 
                            levels = c("Totalmente en desacuerdo", "En desacuerdo",
                                       "Ni de acuerdo ni en desacuerdo", "De acuerdo",
                                       "Muy de acuerdo")))
desc.resp.mmq.long <-   
  desc.resp.mmq.long.pre %>%
  dplyr::count(item,respuesta) %>% group_by(item) %>% 
  mutate(porc = round(n/sum(n),3)*100)

desc.resp.mmq.wide <- 
  desc.resp.mmq.long %>% 
    dplyr::select(item, respuesta, porc) %>%
    pivot_wider(names_from = respuesta, 
              values_from = porc)

print(desc.resp.mmq.wide)


```



### Descriptives per item: Mean, standard deviation, ...

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}
# dataframe for descriptives and cfa
mmq = datos[,7:39]
mmq = na.omit(mmq)

# descriptives
d1 <- psych::describe(mmq, type = 2, skew=TRUE,range=TRUE)
reactivo.mmq <- data.frame(nombres.mmq)
descriptivos.items <- cbind(reactivo.mmq, d1)

print(descriptivos.items)

```



### Confirmatiory Factor Analysis on original scale (Gori et al., 2021)

```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition
model.mmq.or <- '
refv =~ q0007_0001 + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017 + q0007_0018 + q0007_0031 + q0007_0032

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025 + q0007_0026 + 
  q0007_0030

relatv =~ q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'

# fit 
mmq.fit.or <-cfa(model.mmq.or, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(mmq.fit.or, fit.measures = TRUE)

```


### Modification indices analysis
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# modification indices
mi.or <- modindices(mmq.fit.or, sort=TRUE) 
mi.or.1 <- mi.or[mi.or$op == "=~",]

print(mi.or.1)

```

### CFA on modified scale
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# model definition
model.mmq.mod <- '
refv =~ q0007_0001  + q0007_0006 + q0007_0008 + q0007_0010 + q0007_0016 + 
  q0007_0017  

egov =~q0007_0011 + q0007_0022 + q0007_0024 + q0007_0025

relatv =~ q0007_0004 + q0007_0005 + q0007_0014 + q0007_0021 + q0007_0028 

reldiv =~ q0007_0009 + q0007_0012 + q0007_0015 + q0007_0027 + q0007_0033

distrv =~ q0007_0013 + q0007_0019 + q0007_0020 + q0007_0029

emdisv =~ q0007_0002 + q0007_0003 + q0007_0007 + q0007_0023
'

mmq.fit.mod <-cfa(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

summary(mmq.fit.mod, fit.measures = TRUE)

#summary(mmq.fit, standardized=TRUE)
#summary(mmq.fit.mod, fit.measures = TRUE)

#mmq.fit.mod <-cfa_fit_plot(model.mmq.mod, estimator="WLSMV", ordered=names(mmq), std.lv=TRUE, data=mmq)

#fittable.mod <-nice_fit(mmq.fit.mod, nice_table = TRUE)
#flextable::save_as_docx(fittable.mod, path = "fittable_mod.docx")

#loads.mod <- data.frame(inspect(mmq.fit.mod,what="std")$lambda)
#loads.mod <- round(loads.mod,3)
#loads.mod$names <- rownames(loads.mod)

#write.csv(loads.mod, "loads.mod.csv", row.names = F)


```


### Descriptives on modified scale and subscales and correlations
```{r echo=TRUE, message=FALSE, results=FALSE, warning=FALSE}

# means modified scales
mmq$refle.mod <- rowMeans(mmq[,c( 1,6,8,10,16,17)])
mmq$egost.mod = rowMeans(mmq[,c(11,22,24,25)])
mmq$relatt..mod = rowMeans(mmq[,c(4,5,14,21,28)])
mmq$reldis..mod = rowMeans(mmq[,c(9,12,15,27,33)])
mmq$distr.mod = rowMeans(mmq[,c(13,19,20,29)])
mmq$emodis.mod = rowMeans(mmq[,c(2,3,7,23)])

mmq$mmqPositive <- rowMeans(mmq[,c(34,35,36)])
mmq$mmqNegative <- rowMeans(mmq[,c(37,38,39)])


# test Shapiro
s1 <- apply(mmq[,34:41], 2, shapiro.test)[[1]][[2]]
s2 <- apply(mmq[,34:41], 2, shapiro.test)[[2]][[2]]
s3 <- apply(mmq[,34:41], 2, shapiro.test)[[3]][[2]]
s4 <- apply(mmq[,34:41], 2, shapiro.test)[[4]][[2]]
s5 <- apply(mmq[,34:41], 2, shapiro.test)[[5]][[2]]
s6 <- apply(mmq[,34:41], 2, shapiro.test)[[6]][[2]]
s7 <- apply(mmq[,34:41], 2, shapiro.test)[[7]][[2]]
s8 <- apply(mmq[,34:41], 2, shapiro.test)[[8]][[2]]

shapiro.mod <- unlist(c(s1,s2,s3,s4,s5,s6,s7,s8))


# Cronbach's alpha
a1 <- alpha(mmq[,c( 1,6,8,10,16,17)])$total[2] #Reflexivity
a2 <-alpha(mmq[,c(11,22,24,25)])$total[2] #Ego-Strenght
a3 <-alpha(mmq[,c(4,5,14,21,28)])$total[2]#Relational Attunement
a4 <-alpha(mmq[,c(9,12,15,27,33)])$total[2]# Relational Discomfort
a5 <-alpha(mmq[,c(13,19,20,29)])$total[2]#Distrust
a6 <-alpha(mmq[,c(2,3,7,23)])$total[2]#Emotional Discontrol
apos <- alpha(mmq[,c( 1,6,8,10,16,17, 11,22,24,25,4,5,14,21,28)])$total[2]
aneg <- alpha(mmq[,c(9,12,15,27,33,13,19,20,29,2,3,7,23)])$total[2]

alphas.mod <- unlist(c(a1,a2,a3,a4,a5,a6,apos,aneg))


d2 <- psych::describe(mmq[,34:41], type = 2, skew=TRUE,range=TRUE)
descrip.scales <- cbind(d2,shapiro.mod,alphas.mod)
descrip.scales$scales <- rownames(descrip.scales)

print(descrip.scales)


# correlation among scales
rcorr(as.matrix(mmq[,34:39]))


```



### Convergent and divergent analysis: correlation matrix
```{r eval = TRUE, message=FALSE, results=FALSE, warning=FALSE}

# create dataframe with all subscales
# MMQ
scales_mmq = datos %>% dplyr::select(starts_with("q0007"))

scales_mmq$mmq_ref <- rowMeans(scales_mmq[,c( 1,6,8,10,16,17)])
scales_mmq$mmq_ego = rowMeans(scales_mmq[,c(11,22,24,25)])
scales_mmq$mmq_relat = rowMeans(scales_mmq[,c(4,5,14,21,28)])
scales_mmq$mmq_reldi = rowMeans(scales_mmq[,c(9,12,15,27,33)])
scales_mmq$mmq_dist = rowMeans(scales_mmq[,c(13,19,20,29)])
scales_mmq$mmq_emdis = rowMeans(scales_mmq[,c(2,3,7,23)])

scales_mmq$mmqPositive <- rowMeans(scales_mmq[,c(34,35,36)])
scales_mmq$mmqNegative <- rowMeans(scales_mmq[,c(37,38,39)])

scales_mmq <- 
  scales_mmq  %>%
    dplyr::select(mmq_ref, mmq_ego, mmq_relat, mmq_reldi, mmq_dist, mmq_emdis,
                  mmqPositive,mmqNegative)


#Empathy Quotient [EQ]: ec: cognitive empathy; hs: social skills; re: emotional reactivity
scales_eq <- datos %>% dplyr::select(starts_with("q0009"))

scales_eq <- scales_eq %>% 
  dplyr::mutate(across(.cols = c(4,8,12,14,15,27,32,35,46,48,49,50), .fns = function(x){x=5-x}))

scales_eq=scales_eq %>%
  dplyr::mutate(across(.cols = everything(), .fns = function(x){recode(x, `1`= 2, `2`=1, `3`=0, `4`=0)}))

scales_eq$eq_ec <- rowMeans(scales_eq[,c(1,19,25,26,36,41,43,44,52,54,55,58,60)])
scales_eq$eq_hs <- rowMeans(scales_eq[,c(4,8,12,14,21,35,57)])
scales_eq$eq_re <- rowMeans(scales_eq[,c(15,27,32,42,46,48,49,50,59)])

scales_eq$eqTotal <- rowMeans(scales_eq[,c(61,62,63)])

scales_eq <- 
  scales_eq  %>%
    dplyr::select(eq_ec, eq_hs, eq_re, eqTotal)


# Interpersonal Reactivity Index (IRI). fs: fantasy scale; pt: perspective taking; ec :empathic concern; pd: personal distress 
scales_iri = datos %>% dplyr::select(starts_with("q0010"))
scales_iri$fs = NULL
scales_iri = scales_iri %>% dplyr::mutate(across(.cols = c(3,4,7,12,13,14,15,18,19), .fns = function(x){6-x}))
scales_iri = scales_iri %>% dplyr::mutate(across(.cols = everything() , .fns = function(x){x-1}))

scales_iri$iri_fs = rowMeans(scales_iri[,c(1,5,7,12,16,23,26)])
scales_iri$iri_pd = rowMeans(scales_iri[,c(6,10,13,17,19,24,27)])
scales_iri$iri_pt = rowMeans(scales_iri[,c(3,8,11,15,21,25,28)])
scales_iri$iri_ec = rowMeans(scales_iri[,c(2,4,9,14,18,20,22)])

scales_iri$iriTotal <- rowMeans(scales_iri[,c(29,30,31,32)])

scales_iri <- 
  scales_iri  %>%
    dplyr::select(iri_fs, iri_pd, iri_pt, iri_ec, iriTotal)


# Buss Perry Aggression Questionnaire (BPAQ): af: physical agression; av: verbal agression; ang: anger; host: hostility
scales_bpaq = datos %>% dplyr::select(starts_with("q0011"))

scales_bpaq$bpaq_af = rowMeans(scales_bpaq[,c(1,5,9,17,21,27,29)])
scales_bpaq$bpaq_av = rowMeans(scales_bpaq[,c(2,6,14,18)])
scales_bpaq$bpaq_ang = rowMeans(scales_bpaq[,c(3,7,11,22)])
scales_bpaq$bpaq_host = rowMeans(scales_bpaq[,c(4,16,20,26,28)])

scales_bpaq$bpaqTotal = rowMeans(scales_bpaq[,c(30,31,32,33)])

scales_bpaq <- 
  scales_bpaq  %>%
    dplyr::select(bpaq_af, bpaq_av, bpaq_ang, bpaq_host)


datos_cd <- cbind(
scales_mmq, scales_eq, scales_iri, scales_bpaq)


# correlation matrix
rcorr(as.matrix(datos_cd))


```






